---
title: "Markesan District Chromebooks"
author: "By: Victoria Bengry"
output: html_document
date: "2023-09-13"
---
## Introduction

>The dataset for this project was collected between the years of 2020 and 2023 during the summer inventory process for Chromebooks at Markesan District Schools. The data includes the variable of damage level which was assessed for each Chromebook that was used and collected each year. Many factors can impact the damage level of each Chromebook but in this analysis, there is a focus on the school year and age of Chromebook. 

## Background

>The data involved in this project was collected by myself through my multiple years of work with Markesan District Schools. The original intention of this data was to keep current records of the high school Chromebooks in circulation throughout the school year. This data has now gained a greater purpose, as of now the district is looking at students being assigned the same Chromebook throughout their high school career. This way, in the end, the user can keep their device as a commendment of their graduation and the district can eliminate unnecessary technology waste. Chromebooks are not as easily managed when they can no longer update which inevitably happens to Chromebooks as they age, which leads to e-waste. Another potential benefit of this system could be less damaged Chromebooks. If students had a sense of responsibility for their own Chromebook they would ideally want to keep it in good condition. 

>The original data frame for this project contained 928 entries and 5 variables. Each of the variables was kept for analysis and the number of data entries was paired down to 697 when entries with missing data were removed. Entries with missing values were removed to eliminate the risk of involving intermixed data of middle school Chromebooks with high school ones. The five variables included in the data are Chromebook name, damage level, grade level, sex, and school year. 

## Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo = TRUE, message = FALSE, warning = TRUE,
                      error = TRUE, fig.height = 4)
library(tidyverse)
library(scales)
library(lubridate)
library(kableExtra)
```

```{r, include=FALSE}
chromebook_data = read_csv("data/chromebook_data.csv")
```

```{r, include = FALSE} 
chromebook_data = chromebook_data %>% 
  select("Chromebook", "Damage Level", "Grade Level", "Male or Female", "Year", "Chromebook Age") %>% 
  rename(chromebook = "Chromebook",
         damage_level = "Damage Level",
         grade_level = "Grade Level",
         sex = "Male or Female",
         school_year = "Year",
         chromebook_age = "Chromebook Age"
         ) 
```

```{r,  include=FALSE}
chromebook_data_count0 = chromebook_data %>% 
  group_by(school_year) %>% 
  count()
```

> First, I took a look at the total number of Chromebooks collected and put into inventory for each year.

```{r,echo = FALSE}
ggplot(chromebook_data_count0, aes(x = school_year, y = n)) +
  geom_col(fill = "steelblue") +
  labs(title = "Total Chromebooks by Year",
       x = "Year",
       y = "Number of Chromebooks")
```

The number of Chromebooks collected after each school year increases over time which is due to the supply of district chromebooks and possibly more middle school chromebooks being accounted for. 

```{r,include=FALSE}
chromebook_data = na.omit(chromebook_data)
```


```{r,include=FALSE}
chromebook_data_count = chromebook_data %>% 
  group_by(school_year) %>% 
  count()
```

> Here I looked at the total Chromebooks collected with the full information identified.  

```{r,echo = FALSE}
ggplot(chromebook_data_count, aes(x = school_year, y = n)) +
  geom_col(fill = "steelblue") +
  labs(title = "Collected Chromebooks by Year (That were assigned)",
       x = "Year",
       y = "Number of Chromebooks")
```

The total number of Chromebooks collected with full information is about the same throughout the years. Small changes can be accounted for by changing class sizes. 

```{r, include=FALSE}
chromebook_data_damage = chromebook_data %>% 
  group_by(damage_level) %>% 
  count()
```

> Next, I explored the damage level of each Chromebook across all school years. 

```{r, echo = FALSE}
ggplot(chromebook_data_damage, aes(x = damage_level, y = n)) +
  geom_col(fill = "steelblue") +
  labs(title = "Chromebook Damage Level",
       x = "Damage Level",
       y = "Number of Chromebooks")
```

The damage level of Chromebooks throughout the three school years is for the most part "None". 

```{r, include=FALSE}
chromebook_data_damage2 = chromebook_data %>% 
  group_by(damage_level, school_year) %>% 
  filter (damage_level != "None") %>%
  count()
```

> Comparisons between the school years revealed more about the damage levels over time. 

```{r, echo = FALSE}
ggplot(chromebook_data_damage2, aes(x = damage_level, y = n)) +
  geom_col(fill = "steelblue") +
  labs(title = "Chromebook Damage Level (Excluding No Damage)",
       x = "Damage Level",
       y = "Number of Chromebooks") +
  scale_x_discrete(limits = c("Low", "Medium", "High", "Parts")) +
  facet_wrap(~school_year)

```

Damage level excluding the "None" category to better evaluate the rest of the variables. It can be observed that the "Low" and "High" categories are significant across all of the school years. 

```{r,  include=FALSE}
ggplot(chromebook_data_damage2, aes(x = damage_level, y = n)) +
  geom_col(fill = "steelblue") +
  labs(title = "Chromebook Damage Level (Excluding No Damage)",
       x = "Damage Level",
       y = "Number of Chromebooks") +
  scale_x_discrete(limits = c("Low", "Medium", "High", "Parts")) +
  facet_wrap(vars(school_year))
```

```{r,  include=FALSE}
chromebook_data_age= chromebook_data %>% 
  group_by(chromebook_age, school_year) %>% 
  count()
```

> Finally, based on the increasing damage levels across the school years, I evaluated the overall age of the Chromebooks that were given to students. 

```{r,echo = FALSE}
ggplot(chromebook_data, aes(x = school_year, y = chromebook_age)) +
  geom_col(fill = "steelblue") +
  labs(title = "Chromebook Age",
       x = "Year",
       y = "Ages of Chromebooks Combined") 
```

The overall age of the devices can potentially lead to problems as updates become less available. 

## Conclusion

> In the 2020-2021 school year, Chromebooks were taken home by the student and brought back to school each day. In the 2021-2022 school year students were not prohibited to take the Chromebooks home and instead the Chromebooks were left in charging units overnight. In the 2022-2023 school year students were able to take their Chromebooks home with them again. When comparing the damage levels above, it can be seen that the damage level was sustained throughout the first two school years but increased in the 2022-2023 school year. This increase in damage at the end could be due to students taking the Chromebooks home as well as the age of the Chromebooks. The lack of change within the first two school years suggests that the Chromebooks sustained less damage when they were not permitted to be taken home by the students, despite being slightly older. 

> Overall, the data shows that the best course of action is to keep the Chromebooks at school overnight in order to prevent unnecessary damage. As discussed in the background, this conclusion has an impact on the possibility of students having a Chromebook throughout their high school career. Since more damage occurred when the students were permitted to carry their Chromebooks beyond the school day, this does not seem to be an ideal situation. Given this information, I believe there is another solution that allows the student to take responsibility, reduce electronic waste, and reduce overall damage caused by Chromebooks. If students were only permitted to use the Chromebooks at school but were still provided with a Chromebook after graduation they would likely take better care of the Chromebooks and there would be less waste overall. As more data is collected, these results could be further explored. 
